---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# GWPR.light

<!-- badges: start -->
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![Travis build status](https://travis-ci.com/MichaelChaoLi-cpu/GWPR.light.svg?branch=master)](https://travis-ci.com/MichaelChaoLi-cpu/GWPR.light)
[![R-CMD-check](https://github.com/MichaelChaoLi-cpu/GWPR.light/workflows/R-CMD-check/badge.svg)](https://github.com/MichaelChaoLi-cpu/GWPR.light/actions)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/MichaelChaoLi-cpu/GWPR.light?branch=master&svg=true)](https://ci.appveyor.com/project/MichaelChaoLi-cpu/GWPR.light)
<!-- badges: end -->

This package includes the function for the optimal bandwidth selection in GWPR, the function for GWPR, the function for the local Hausman test, the function for the local F test for individual effects, the function for the local Lagrange Multiplier Breusch-Pagan test, and the function for panel Moran's I test. The functions have been optimized, which require the less memory in the calculation. 

## Author 

Chao Li <chaoli0394@gmail.com> Shunsuke Managi <managi@doc.kyushu-u.ac.jp>

## Maintainer

Chao Li <chaoli0394@gmail.com>


## Installation

You can install the released version of GWPR.light from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("GWPR.light")
```

## Detailed Introduction
You can read our [vignettes](/vignettes/introduction_of_GWPR.html)

## Example

This is a basic example which shows you how to solve a common problem:
```{r, message=FALSE}
library(GWPR.light)
```
```{r example}
library(tmap)
## basic example code
data(TransAirPolCalif)
data(California)
formula.GWPR <- pm25 ~ co2_mean + Developed_Open_Space_perc + Developed_Low_Intensity_perc +
   Developed_Medium_Intensity_perc + Developed_High_Intensity_perc +
   Open_Water_perc + Woody_Wetlands_perc + Emergent_Herbaceous_Wetlands_perc +
   Deciduous_Forest_perc + Evergreen_Forest_perc + Mixed_Forest_perc +
   Shrub_perc + Grassland_perc + Pasture_perc + Cultivated_Crops_perc +
   pop_density + summer_tmmx + winter_tmmx + summer_rmax + winter_rmax
```

This is an example about GWPR.moran.test:
```{r eg.GWPR.moran.test}
pdata <- plm::pdata.frame(TransAirPolCalif, index = c("GEOID", "year"))
moran.plm.model <- plm::plm(formula = formula.GWPR, data = pdata, model = "within")
#summary(moran.plm.model)

bw.AIC.F <- bw.GWPR(formula = formula.GWPR, data = TransAirPolCalif, index = c("GEOID", "year"), SDF = California,
                     adaptive = F, p = 2, bigdata = F, effect = "individual",
                     model = "within", approach = "AIC", kernel = "bisquare", longlat = F,
                     doParallel = T, cluster.number = 4)

# moran's I test
GWPR.moran.test(moran.plm.model, SDF = California, bw = bw.AIC.F, kernel = "bisquare",
                 adaptive = F, p = 2, longlat=F, alternative = "greater")
# The statistic is significantly greater than 0. Therefore, the residuals are spatially clustered.
```

GWPR example:

```{r eg.GWPR}
result.F.AIC <- GWPR(bw = bw.AIC.F, formula = formula.GWPR, data = TransAirPolCalif, index = c("GEOID", "year"),
                     SDF = California, adaptive = F, p = 2, effect = "individual", model = "within",
                     kernel = "bisquare", longlat = F)
summary(result.F.AIC$SDF$Local_R2)
tm_shape(result.F.AIC$SDF) +
  tm_polygons(col = "Local_R2", pal = "Reds",auto.palette.mapping = F,
              style = 'cont')
```

Example of F test:
```{r eg.GWPR.pFtest}
GWPR.pFtest.resu.F <- GWPR.pFtest(formula = formula.GWPR, data = TransAirPolCalif, index = c("GEOID", "year"),
                                  SDF = California, bw = bw.AIC.F, adaptive = F, p = 2, effect = "individual",
                                  kernel = "bisquare", longlat = F)
tm_shape(GWPR.pFtest.resu.F$SDF) +
     tm_polygons(col = "p.value", breaks = c(0, 0.05, 1))
```

Example of Locally Breusch-Pagan Lagrange Multiplier Test:
```{r eg.GWPR.plmtest}
GWPR.plmtest.resu.F <- GWPR.plmtest(formula = formula.GWPR, data = TransAirPolCalif, index = c("GEOID", "year"),
                                    SDF = California, bw = bw.AIC.F, adaptive = F, p = 2,
                                    kernel = "bisquare", longlat = F)
tm_shape(GWPR.plmtest.resu.F$SDF) +
     tm_polygons(col = "p.value", breaks = c(0, 0.05, 1))
```

Example of Locally Hausman Test Based on GWPR
```{r eg.GWPR.phtest}
GWPR.phtest.resu.F <- GWPR.phtest(formula = formula.GWPR, data = TransAirPolCalif, index = c("GEOID", "year"),
                                  SDF = California, bw = bw.AIC.F, adaptive = F, p = 2, effect = "individual",
                                  kernel = "bisquare", longlat = F, random.method = "amemiya")
tm_shape(GWPR.phtest.resu.F$SDF) +
     tm_polygons(col = "p.value", breaks = c(0, 0.05, 1))
```

